---
title: "nflspread"
author: "JMeerse"
date: "2022-09-05"
output: html_document
---
## Home field advantage in the NFL

We'll explore 10 years of data from 2012 - 2021 to see if it exists, and what the advantage is.

```{r setup, results="hide"}

# install.packages("nflreadr")
# install.packages("nflfastR")
# install.packages("nflplotR")

library(nflreadr)
library(tidyverse)
library(nflfastR)
library(ggimage)
library(nflplotR)
```

# Load data
The nflreadR package access to the weekly results for each season loaded.
```{r}
seasons <- load_schedules(seasons =2012:2021)
```

Use the data dictionary to see what each variable represents.

<https://nflreadr.nflverse.com/articles/dictionary_schedules.html>

# Looking at the data.

We need a new variable to determine if the home or away team won the game.

`mutate()` creates a new variable.

`if_else()` works as: 
   "check condition", 
  "if true assign first value", 
  "if false assign second value."
```{r}
seasons <- seasons %>% 
  mutate(winner = if_else(result > 0, home_team, away_team),
         loser = if_else(result < 0, home_team, away_team))

seasons <- seasons %>% 
  mutate(win_ha = if_else(winner == home_team, "home", "away"))

seasons <- seasons[seasons$result != 0, ] #this gets rid of ties
```

## Regular season vs Playoffs
We'll create 2 more dataframes, one for each season type.
```{r}
seasons_reg <- seasons %>% filter(game_type == "REG")
seasons_plyoff <- seasons %>% filter(game_type != "REG")

```

## Now lets see how home teams perform.
```{r}
seasons_reg %>% ggplot(aes(x = win_ha)) + geom_bar() #bar chart of overall home or away winning percentage

seasons_reg %>% count(win_ha) %>% mutate(prop = n / sum(n)) # table of winning percents for entire 10 year period

```

## Let's break things down by season
```{r}
by_season <- seasons_reg %>% group_by(season, win_ha) %>%
  summarise(count = n()) %>%
  mutate(freq = round(count / sum(count), 3)) %>% ungroup()

by_season

by_season %>% filter(win_ha == "home") %>% 
  ggplot(aes(x = season, y = freq)) + 
  geom_line() +
  labs()

#repeat the above for the playoffs.  
#Can we group_by home team to see the effect at each stadium? 
```
